---
title: "Getting started with PhysioMSKNet"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with PhysioMSKNet}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 6, fig.height = 4)
library(PhysioMSKNet)
set.seed(1)
```

PhysioMSKNet treats the musculoskeletal system as a **hypergraph**: bones are
vertices and muscles are hyperedges connecting the bones they attach to
(Murphy et al. 2018). This vignette walks through the core workflow — build a
hypergraph, measure it, detect communities, score perturbation impact, and
compare impact against a degree-preserving null model — using a small synthetic
network first and then the bundled dataset. Everything here runs offline.

## A hypergraph from an incidence matrix

The incidence matrix `C` has one row per bone and one column per muscle, with
`C[i, j] = 1` when muscle *j* attaches to bone *i*. Here is a tiny four-bone,
four-muscle ring:

```{r build}
C <- matrix(c(1, 1, 0, 0,   0, 1, 1, 0,   0, 0, 1, 1,   1, 0, 0, 1),
            nrow = 4, dimnames = list(paste0("bone", 1:4), paste0("m", 1:4)))
hg <- MSKHypergraph(C)
hg
```

The degree of a muscle (hyperedge) is the number of bones it spans; the degree
of a bone (vertex) is the number of muscles attached to it:

```{r degrees}
hyperedgeDegree(hg)
vertexDegree(hg)
```

## Projections and network metrics

Collapsing the hypergraph onto the bones gives a weighted bone–bone graph; the
network metrics are computed on that projection.

```{r metrics}
as.matrix(projectBoneGraph(hg))
m <- mskNetworkMetrics(hg, type = "bone")
m$density
m$degree
```

## Community detection

`mskCommunityDetect()` runs Louvain modularity with a resolution parameter
`gamma`. On the bundled network the paper's resolution (`gamma = 4.3`) yields
the functional muscle communities:

```{r community}
hg_full <- MSKHypergraph()
comm <- mskCommunityDetect(hg_full, gamma = 4.3)
comm$n_communities
plotCommunityStructure(comm)
```

## Perturbation impact

The impact score perturbs one muscle in a damped harmonic-oscillator model and
sums the resulting displacement across all bones. `mskSimulate()` sets the
model up; a short run is enough for the synthetic network:

```{r impact}
sim <- mskSimulate(hg, n_steps = 100)
scores <- mskImpactScoreAll(sim, verbose = FALSE)
scores
```

## Comparison against a null model

To tell a structurally important muscle from one that merely has a high degree,
compare its impact against a degree-preserving null ensemble. The deviation is
expressed in standard deviations of the null distribution:

```{r null}
ens <- mskNullEnsemble(hg, n_null = 20, sim_params = list(n_steps = 100),
                       verbose = FALSE)
mskImpactDeviation(scores, hg, ens$null_scores)
```

## Reproducing the paper's statistics

The bundled validation tables let you reproduce the headline relationships
without re-running the full simulation. For example, the impact-deviation /
clinical-recovery regression (Fig. 3b):

```{r reproduce}
rec <- mskImpactRecoveryModel()
rec$r_squared
rec$paper_target$r_squared
```

## Where to go next

- Bridge the network to measured signals with `emgToMSKMapping()`,
  `imuToMSKMapping()` and `mocapToMSKMapping()`, then the matched analyses
  (`imuNetworkKinematics()`, `imuImpactPrediction()`, ...).
- The EMG-driven Hill muscle model (`emgDrivenForce()`,
  `calibrateEMGDrivenModel()`) turns an EMG envelope into muscle force and joint
  moment.
- See `?PhysioMSKNet` for the full, task-grouped index of entry points.

## Reference

Murphy AC, Muldoon SF, Baker D, Lastowka A, Bennett B, Yang M, Bhatt P,
Bassett DS (2018). "Structure, function, and control of the human
musculoskeletal network." *PLOS Biology* 16(1): e2002811.
